---
title: "Hanukkah of Data - Day 3"
author: "Clark Seanor"
format: gfm
  # html:
    # df-print: paged
    # code-fold: true
---

## The question

The objective of today's challenge is to find a person's phone number who was born in the year of the dog and is an Aries.


## My approach

### Opening up the data 
First, I want to have a quick look at the customers table.

```{r}
#| warning: false

library(readr)
library(dplyr)

customers = read_csv("data/noahs-customers.csv")
glimpse(customers)
```

I'm not totally sure why the birthdate is stored here, but whatever?

Anyway, because the date comparison library I wanted to use doesn't seem to work properly on my computer because it can't talk to the time management system or whatever ***(I blame Crouton)*** I'm just going to split out the date parts I want to be able to use. Aries are born between March 21 and April 20, according to Google. The year of the dog occurs every 12 years, when `year % 12 = 2`. So I need to be able to use all these date parts, and I need to be able to use them as ints.

```{r}
customers <- customers %>%
 select(name, phone, birthdate, customerid) %>%
 mutate(
  year = as.integer(format(customers$birthdate, "%Y")),
  month = as.integer(format(customers$birthdate, "%m")),
  day = as.integer(format(customers$birthdate, "%d")),
 )

customers <- customers[customers$year %% 12 == 2,]

customers <- customers[(customers$month == 3 & customers$day > 20) | (customers$month == 4 & customers$day < 21),]

glimpse(customers)

```

So it turns out that there are a lot of Aries born in the year of the dog. Okay. So we also know that the person probably visited Noah's in 2017.

```{r}
#| warning: false
orders = read_csv("data/noahs-orders.csv")
orders <- orders[format(orders$ordered, "%Y") == 2017,]
customers <- merge(x=customers, y=orders, by="customerid")

unique(customers$name)
```

Well. There are still far too many of them. So maybe one of them bought rug cleaner (SKU HOM8601)?


```{r}
#| warning: false
orders_items = read_csv("data/noahs-orders_items.csv")
orders_items <- orders_items[orders_items$sku == "HOM8601",]
customers <- merge(x=customers, y=orders_items, by="orderid")

customers
```

Well... since it's a guy it's probably either going to be Carlos or Brent. Since Brent bought the most rug cleaner and the rug was really dirty, I decided to enter Brent's phone number first. It was Brent :)
